---
title: "Chicago Food Inspections"
output: html_notebook
---

```{r setup}
library(readr)
library(dplyr)



readrestaurant<-function(x){
  foodinspectioncols<-c(
      "InspectionID",	"DBAName",	"AKAName",	"LicenseID",	"FacilityType", "Risk"
      , "Address",	"City",	"State",	"Zip"
      , "InspectionDate","InspectionType",	"InspectionResults", "InspectionNotes"
      ,	"Latitude",	"Longitude", "LatLong"
      )
  read_csv(x
           , col_names = foodinspectioncols
           , col_types = cols("InspectionDate"=col_date(format = "%m/%d/%Y"))
           )
}
p1<-readrestaurant("../FoodInspection/Food_Inspections1.csv")
p2<-readrestaurant("../FoodInspection/Food_Inspections2.csv")

inspections<-union(p1,p2)
```

# Data exploration
```{r}
inspections %>%
  summary
```


## Handling potential factors

```{r}
inspections %>%
  select_if(is.character) %>%
  summarise_each(funs(uniqueN=n_distinct))
```

### Facility Type
```{r}
inspections$FacilityType<-tolower(inspections$FacilityType)
(inspections %>%
   mutate(total=n()) %>%
   group_by(FacilityType) %>%
   summarise(N=n(), total=min(total)) %>%
   mutate(P=N/total) %>%
   arrange(desc(N)) %>%
   filter(!is.na(FacilityType))->
   facilities)
```

Perhaps we can combine these? Instead of doing it manually, perhaps we can do some sort of fuzzy grouping?

```{r}
library(stringdist)
library(forcats)
distances <- stringdistmatrix(facilities$FacilityType, facilities$FacilityType)
distances[1:10,1:10]
weightings <- 10^facilities$P#1#log(facilities$N+1)
weightings[1:10]
weighteddistances <- distances / weightings
weighteddistances[c(1,60),c(1,60)]
diag(weighteddistances) <- 9999999
matches <- apply(weighteddistances,2, which.min)
matches[1:10]
facilities %>%
  mutate(NextClosest= FacilityType[matches]) %>%
  mutate(Group= ifelse(P>=0.05,FacilityType,NextClosest)) ->
  facilities

t(facilities[1:7,])
```

Looking at the revised groups
```{r}
facilities %>%
  group_by(Group) %>%
  summarise(N=sum(N)) %>%
  arrange(desc(N)) ->
  groupedfacilities

groupedfacilities
```

There's still a long tail but it's much better! Let's apply these new values

```{r}
inspections %>%
  left_join(facilities) %>%
  select(-N, -P, -total, -NextClosest, RefinedFacilityType=Group) %>%
  mutate(RefinedFacilityType=factor(RefinedFacilityType)) ->
  inspections

inspections %>%
  group_by(RefinedFacilityType) %>%
  summarise(N=n())
```

We can then use forcats to consolidate the long tail
```{r}
inspections %>%
  mutate(RefinedFacilityType= fct_lump(RefinedFacilityType, prop=0.01)) ->
  inspections
```

### Risk
```{r}
inspections %>%
  group_by(Risk) %>%
  summarise(N=n())
```

```{r}
inspections %>%
  mutate(Risk = ordered(Risk, levels = c("Risk 3 (Low)", "Risk 2 (Medium)","Risk 1 (High)"), labels = c("L", "M","H"))) ->
  inspections
```

### City
```{r}
inspections %>%
  group_by(City) %>%
  summarise(N=n())
```

This doesn't look like a field of interest. Let's quickly check state too.

```{r}
inspections %>%
  group_by(State) %>%
  summarise(N=n())
```
One missing, and as expected Chicago is in Illinois!

### Inspection Types

```{r}
inspections %>%
  group_by(InspectionType) %>%
  summarise(N=n())
```

```{r}
library(stringr)
inspections %>%
  mutate(RefinedInspectionType = str_to_lower(InspectionType)) %>%
  mutate(RefinedInspectionType = word(RefinedInspectionType, 1, sep = " |-|/")) %>%
  mutate(RefinedInspectionType = fct_infreq(RefinedInspectionType)) %>%
  mutate(RefinedInspectionType = fct_lump(RefinedInspectionType, prop=.05)) %>%
  mutate(RefinedInspectionType = fct_collapse(RefinedInspectionType, complaint="short")) ->
  inspections

```

```{r}
inspections %>%
  group_by(RefinedInspectionType) %>%
  summarise(N=n())
```


### Pass
```{r}
inspections %>%
  group_by(InspectionResults) %>%
  summarise(N=n())
```

```{r}
inspections %>%
  mutate(RefinedInspectionResults = fct_collapse(
    InspectionResults, Other = c("Business Not Located","Out of Business"))) ->
  inspections
```

```{r}
inspections %>%
  group_by(RefinedInspectionResults) %>%
  summarise(N=n())
```

### Inspection notes
```{r}
inspections %>%
  select(InspectionNotes) %>%
  head
```

It looks like we could be able to split these out by the pipe(`|`) symbol. 

```{r}
library(data.table)
library(dtplyr)
library(dplyr)

inspections %>%
  select(InspectionID, InspectionNotes) %>%
  setDT -> 
  notes

notes<-cbind(notes, notes[,tstrsplit(InspectionNotes,split=" | ", fixed=TRUE)])

notes %>%
  select(-InspectionNotes) %>%
  melt(id.var=c("InspectionID"), na.rm=TRUE) %>%
  select(-variable) %>%
  mutate(note = as.numeric(word(value,sep = fixed(".")))) ->
  notespivot

inspections %>%
  setDT %>%
  full_join(notespivot) ->
  notesinspections
```

```{r}
notesinspections %>%
  group_by(note, RefinedInspectionResults) %>%
  summarise (N=n()) ->
  notefrequencies

notefrequencies[note==30,]
notespivot[note==30,]
```

```{r}
library(ggplot2)
ggplot(notefrequencies, aes(x=note, y=N
                            , group=RefinedInspectionResults
                            , fill=RefinedInspectionResults)) +
  geom_col()
```

```{r}
ggplot(notefrequencies, aes(x=note, y=N
                            , group=RefinedInspectionResults
                            , fill=RefinedInspectionResults)) +
  geom_col(position="fill")
```

# Pass rates

## By Facility Type

```{r}
library(ggplot2)

ggplot(inspections,aes(x=RefinedFacilityType, group=RefinedInspectionResults, fill=RefinedInspectionResults))+
  geom_bar()+ coord_flip()
```

## By Inspection Type

```{r}
ggplot(inspections,aes(x=RefinedInspectionType, group=RefinedInspectionResults, fill=RefinedInspectionResults))+
  geom_bar()+ coord_flip()
```

# Geography
```{r}
library(ggmap)
ggmap(get_map(location = 'Chicago', zoom = 11)) +
  geom_point(data=inspections, aes(x=Longitude, y=Latitude, colour=RefinedInspectionResults), alpha=.2)
```

```{r}
library(ggmap)
ggmap(get_map(location = 'Chicago', zoom = 11)) +
  geom_density2d(data=inspections, aes(x=Longitude, y=Latitude, colour=RefinedInspectionResults))
```
