---
title: "Time Series basics"
output: html_notebook
---

# Time series packages

- **stats** base R functionality
    + `ts()`
- **forecast** forecasting functionality
- **zoo** and **xts** improved time series manipulations
- **ggseas** ggplot2 time series
- **fpp** practice datasets

Get all (good) time series packages via the CRAN task view:

```r
library(ctv)
install.views("TimeSeries")
```

```{r}
library(forecast)
library(zoo)
library(ggseas)
library(ggplot2)
library(fpp)
```

# Time series fundamentals

## Making time series
Time series need:
- **values** the thing being measured
- **interval** typically this is constant
- **frequency** number of records in a cycle

### From vector
```{r}
kings <- scan("http://robjhyndman.com/tsdldata/misc/kings.dat",skip=3)
ts(kings)
```

### From data.frame
```{r}
sample_df<-data.frame(l=seq(Sys.Date(),by = 1, length.out = 26), r=rnorm(26))
smpts<-zoo(sample_df$r,order.by = sample_df$l)
smpts
```

## Basic plotting

### Base R
```{r}
plot(ts(kings))
```

### ggplot2
```{r}
# via forecast funcction autpolot
autoplot(smpts)
```

## Decomposition
```{r}
# via ggseas
ggsdc(tsdf(usconsumption), aes(x=x, y=consumption), s.window=4)+geom_line()
```

# Moving averages
Use moving averages to remove noise from a timeseries

```{r}
cons<-zoo(usconsumption)
zmean<-rollmean(cons,k=4)
zmedian<-rollmedian(cons,k=4)
results<-merge(merge(cons,zmean),zmedian)
autoplot(results)
```

# Smoothing
Build a curve that can be extended for forecasts

```{r}
etscons<-ets(cons$consumption)
autoplot(etscons)
```

```{r}
forecons<-forecast(etscons, h=4)
autoplot(forecons)
```

```{r}
accuracy(forecons)
```

# ARIMA
Take into account correlations between succesive values of a stationary time series.

>Use the log of a timeseries if it's multiplicative or take a diff between successive values

```{r}
diffs<-diff(cons)
consanddiffs<-merge(cons,diffs)
autoplot(consanddiffs)
```

```{r}
acf(diffs)
```

```{r}
pacf(diffs)
```

[auto.arima alogirithm](https://www.otexts.org/fpp/8/7)
```{r}
diffsarima<-auto.arima(diffs$consumption)
diffsarima
```

```{r}
autoplot(forecast(diffsarima,h=4))
```

```{r}
consarima<-auto.arima(cons$consumption)
consarima
```

```{r}
autoplot(forecast(consarima,h=4))
```

```{r}
autoplot(consarima)
```

```{r}
accuracy(consarima)
```

# Regression forecasts
For a multivariate forecast

```{r}
autoplot(cons)
qplot(cons$income, cons$consumption)
```

```{r}
lmusc <- tslm(consumption ~ income, data=usconsumption)

qplot(cons$income, cons$consumption) + geom_abline(intercept=coef(lmusc)[1],slope=coef(lmusc)[2])
```

```{r}
scenarios<-forecast(lmusc, newdata=data.frame(income=-1:1))

autoplot(scenarios)
```

```{r}
autoplot(ts(resid(lmusc)))
```

```{r}
Acf(ts(resid(lmusc)))
```

```{r}
accuracy(lmusc)
```